---
title: "Supplementary Material"
header-includes:
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{titlesec}
  - \usepackage{longtable}  # Ensure longtable is included for better table formatting
  - \usepackage{booktabs}   # For better quality table rules
  - \newcommand{\beginsection}{\clearpage}
  - \renewcommand{\thefigure}{S\arabic{figure}}
  - \renewcommand{\thetable}{S\arabic{table}}
  - \titleformat{\section}{\normalfont\Large\bfseries}{\thesection}{1em}{}[\vspace{-0.5\baselineskip}]
output:
  pdf_document:
    fig_width: 6.5
    fig_height: 7.4
    keep_tex: yes  # Keep the LaTeX file for troubleshooting
    extra_dependencies: ["longtable", "booktabs"]  # Ensure dependencies are recognized
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 6.5, fig.height = 7.4, fig.align = 'center')
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(cowplot)
library(ggpubr)
library(here)
library(caret)
library(leaps)
library(knitr)

# Adjust these paths to point to your R files
source(here("supplementary", "plot_case.R"))

## Input
## Dataset
dataset <- read_csv(here("data", "final_dataset.csv")) %>%
  mutate(mcsa_id = as.character(mcsa_id))

## Profiles
profiles <- read_csv(here("data", "final_dataset_profiles.csv")) %>%
  mutate(mcsa_id = as.character(mcsa_id))

## Goodness of fit of scam models
data_gof <- read_csv(here("data", "final_dataset_gof.csv")) %>%
  mutate(mcsa_id = as.character(mcsa_id))

## Some dataset entry properties
dataset_table <- read_csv(here("data", "final_dataset_table.csv"))

```


This file contains supplementary tables and figures associated to the paper \emph{On the variation of structural divergence among residues in enzyme evolution}, by J. Echave and M. Carpentier.

\beginsection
# Dataset 
\nopagebreak


```{r, eval = TRUE}
library(kableExtra)

dataset_table %>% 
  mutate(across(where(is.double), ~ round(., 2))) %>% 
  mutate(seqid_mean = round(seqid_mean * 100, 0)) %>% 
  select(mcsa_id, pdb_chain_id, name, source, family) %>% 
  kable(
    col.names = c("M-CSA", "PDB", "Name", "Source", "Family"),
    caption = "Dataset entries and reference protein",
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>% 
  kable_styling(
    font_size = 8,
    latex_options = c("scale_down", "hold_position")  # Using hold_position instead
  ) %>%
  column_spec(3:5, width = "4cm")
```

\clearpage
```{r}
library(kableExtra)
# Second table - structural and classification information
dataset_table %>% 
  mutate(across(where(is.double), ~ round(., 2))) %>% 
  mutate(seqid_mean = round(seqid_mean * 100, 0)) %>% 
  select(mcsa_id, nprot, seqid_mean, rmsd_mean, nresidues, nresidues_asite, rmsd_asite, cath_class, ec_class) %>% 
  kable(
    col.names = c("M-CSA", "N", "<Id%>", "<RMSD>", "Length", "AS size", "AS RMSD", "CATH", "EC"),
    caption = "Dataset properties",
    format = "latex",
    booktabs = TRUE
  ) %>% 
  kable_styling(
    font_size = 8,
    latex_options = "scale_down"
  ) %>% 
  footnote(
    general = "N: number of family members; <Id%>: average family seq. identity; <RMSD>: average family RMSD; Length: number of residues of reference protein; AS size: number of residues of active site; CATH: CATH class; EC: EC class",
    general_title = "",
    threeparttable = TRUE
  )
```

\clearpage



```{r load_data_and_create_plot, fig.width=6.5, fig.height=9, fig.cap="Distribution of properties of dataset families."}
# Create and display the plot
source(here("supplementary", "plot_dataset_distributions.R"))
plot <- plot_dataset_distributions(dataset)
print(plot)
```

\clearpage

\beginsection

# Correlation between flexibility and distance

The following figure shows the distribution of spearman correlations between flexibility (lRMSF) and distance to the active site (d) (panel a).
The correlation between these variables depends on the location of the active site within the enzyme structure. In most cases active sites are located in relatively rigid regions. Therefore, as one moves away from the active site, not only distance increases but flexibility too, which originates the correlation  (panel b).

```{r flex-dist correlation, fig.width=6.5, fig.height=5.5, fig.cap="Correlation betwen flexibility and distance."}
# First calculate the statistics
stats_summary <- profiles %>% 
  group_by(mcsa_id) %>% 
  summarise(rho_rmsf_dactive = cor(lrmsf, dactive, method = "spearman")) %>% 
  summarise(
    mean_rho = mean(rho_rmsf_dactive, na.rm = TRUE),
    sd_rho = sd(rho_rmsf_dactive, na.rm = TRUE),
    min_rho = min(rho_rmsf_dactive, na.rm = TRUE),
    max_rho = max(rho_rmsf_dactive, na.rm = TRUE)
  ) 

# Create the plot with annotations
p1 <- profiles %>% 
  group_by(mcsa_id) %>% 
  summarise(rho_rmsf_dactive = cor(lrmsf, dactive, method = "spearman")) %>% 
  ggplot(aes(rho_rmsf_dactive)) +
  geom_histogram(fill = "grey", color = "black", bins = 10) +
  xlab(bquote(rho(nlRMSF, d))) +
  ylab("Number of cases") +
  annotate(
    "text", 
    x = -Inf, 
    y = Inf,
    hjust = -0.1,
    vjust = 1.5,
    label = sprintf(
      "Range: [%.2f, %.2f]\nMean: %.2f\nSD: %.2f",
      stats_summary$min_rho,
      stats_summary$max_rho,
      stats_summary$mean_rho,
      stats_summary$sd_rho
    )
  )  +
  theme_cowplot()

p2 <- profiles %>% 
  group_by(mcsa_id) %>% 
  summarise(rho_rmsf_dactive = cor(lrmsf, dactive, method = "spearman"),
            nlrmsf_asite = mean(nlrmsf[near(dactive, 0)])) %>% 
  ggplot(aes(nlrmsf_asite, rho_rmsf_dactive)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor() +
  ylim(-0.1, 1) +
  xlab("Active site flexibility (nlRMSF)") +
  ylab(bquote(rho(nlRMSF, d)))  +
  theme_cowplot()

plot_grid(p1, p2, ncol = 1, 
         labels = c("(a)", "(b)"),
         hjust = 0,
         vjust = 0) +    
  theme(plot.margin = unit(c(0.25, 0.25, 0.25, 0.25), "inches"))
```


\clearpage

\beginsection

# Family by family analysis

## Figure Description

Each figure in this section represents a structural divergence analysis for a specific enzyme family, identified by its MCSA ID. 

Each figure is organized as follows:

Observed patterns:
(a) Residue-specific structural divergence profile, showing normalized log-RMSD (nlRMSD) across residues.
(b) Relationship between structural divergence (nlRMSD) and residue flexibility (nlRMSF).
(c) Relationship between structural divergence (nlRMSD) and distance from the active site (d).

Model predictions vs. observations:
(d) Comparison of observed nlRMSD profile with predictions from models M1, M2, and M12.
(e) Flexibility trends: observed data vs. model predictions for nlRMSD vs. nlRMSF.
(f) Distance trends: observed data vs. model predictions for nlRMSD vs. d.

Decomposition into flexibility and distance contributions:
(g) Profile of structural divergence decomposed into flexibility (s1) and distance (s2) components.
(h) Flexibility (s1) component of structural divergence.
(i) Distance (s2) component of structural divergence.

```{r generate_figures, results='asis', fig.show='asis', fig.width=6.5, fig.height=9}
# List of all mcsa_ids to process
d <- data_gof %>% 
  arrange(as.numeric(mcsa_id))
mcsa_ids <- d$mcsa_id
# mcsa_ids <- mcsa_ids[1:1]  # Adjust this if you want to process more or all cases

for (id in mcsa_ids) {
 cat("\n\n\\clearpage\n")
 cat("\\begin{figure}[H]\n\\centering\n")
  
 # Generate the plot for this mcsa_id
 plot <- plot_case(profiles, data_gof, id)
  
 # Print the plot
 print(plot)
 
 # Get the PDB chain ID for this MCSA ID
 pdb_chain_id <- dataset %>%
   filter(mcsa_id == id) %>%
   pull(pdb_chain_id)
 
 # Escape the underscore in the PDB chain ID for LaTeX
 pdb_chain_id_escaped <- gsub("_", "\\\\_", pdb_chain_id)
 
 # Add a caption with MCSA ID and escaped PDB chain ID
 cat(sprintf("\\caption{Structural divergence analysis for enzyme family MCSA ID: %s. Reference protein PDB ID: %s.}\n", id, pdb_chain_id_escaped))
 cat("\\end{figure}\n")
}
```
